---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

if (sum(installed.packages()[,1] %in% "pacman") != 1){
  
  install.packages("pacman")
  
}

pacman::p_load(here)

```

# cloutier

<!-- badges: start -->
[![R-CMD-check](https://github.com/WB-PIDA-Data-Science-Shop/cloutier/blob/master/.github/workflows/Rcmdcheck_overkill.yml)
<!-- badges: end -->

The goal of cloutier is to curate data from the Gallup, VDEM and the VoG towards creating a social contract factsheet
flexdashboard. This is a product of the Public Institutions Data and Analytics Unit within the Institutions Department upon the request of Mathieu Cloutier both of the World Bank.    

## Installation

You can install the development version of cloutier from Github as follows:

```{r}

# install.packages("remotes")
remotes::install_github("WB-PIDA-Data-Science-Shop/cloutier")


```

The package can be loaded as follows: 
```{r example}
library(cloutier)
## basic example code
```

You may reproduce all published datasets within this package by simply this file: 

```{r}

source(here("data-raw/prepare_data.R"))

```


### country level data
```{r}

head(cloutier_dt[cloutier_dt$year >= 2005,])

```

### regional comparators dataset
```{r}

head(regcomp_dt[regcomp_dt$year >= 2005,])

```

### global comparators dataset
```{r}

head(globalcomp_dt[globalcomp_dt$year >= 2005,])

```


